Two applications of logic programming to Coq

### 2021-03-04 papers/types2020/revisions/sec3.tex
3.3. talks abiyt conversion. More todos, a fix to fig 3, new file for dependent prolog tactic

The following
example shows how we can use the above tactic to do FPC-driven logic
programming in Coq and return a Coq proof-term:
\begin{lstlisting}
Inductive insert (x:nat) : list nat -> list nat -> Prop :=
| i_n : is_nat x -> insert x [] [x]
| i_s : forall y: nat, forall ys, x <= y -> insert x (y :: ys) (x :: y :: ys)
| i_c : forall y: nat, forall ys rs, y <= x -> insert x ys rs -> insert x (y :: ys) (x :: rs).
Lemma i1:  exists R, insert 2 [0;1] R.
eexists.
elpi dprolog height 10.
Qed.
Print i1.
ex_intro (fun R : list nat => insert 2 [0; 1] R) [0; 1; 2]
  (i_c 2 0 [1] [1; 2] (i_c 2 1 [] [2] (i_n 2)) : exists R : list nat, insert 2 [0; 1] R
\end{lstlisting}

The \lsti{dprolog} tactic can be seen as a programmable (and possibly
faster) version of Coq's \lsti{eauto}, although it currently lacks the
latter's capability of solving goals by
\lsti{reflexivity}; % as well as the modularity of using
% \lsti{Hints} as introduction rules
 \lsti{dprolog} is not restricted to depth-first search, since
it follows the dictate of the given FPC; for example we could easily add
iterative-deepening search. Furthermore, the tactic can still use
\lsti{Hints} via Elpi's \lsti{Db} feature, and FPCs can provide a
\emph{trace} that may be more customizable than the one offered by 
\lsti{(e)auto}'s hard-wired \lsti{Debug} facility.

>>>>

The following
example shows how we can use the above tactic to do FPC-driven logic
programming in Coq and return a Coq proof-term:
\begin{lstlisting}[]
Inductive insert (x:nat) : list nat -> list nat -> Prop :=
| i_n : insert x [] [x]
| i_s : forall y: nat, forall ys, x <= y -> insert x (y :: ys) (x :: y :: ys)
| i_c : forall y: nat, forall ys rs, y <= x -> insert x ys rs -> insert x (y :: ys) (x :: rs).
Lemma i1:  exists R, insert 2 ([0] $++$ [1]) R.
elpi dprolog height 10.
Qed.
Print i1.
ex_intro (fun R : list nat => insert 2 ([0] $++$ [1]) R)  [0; 1; 2]
  (i_c 2 0 [1] [1; 2] (i_c 2 1 [] [2] (i_n 2))) : exists R : list nat, insert 2 ([0] $++$ [1]) R
\end{lstlisting}

The \lsti{dprolog} tactic can be seen as a programmable (and possibly
faster) version of Coq's \lsti{eauto}% , although it currently lacks the
% latter's capability of solving goals by
% \lsti{reflexivity}
; % as well as the modularity of using
% \lsti{Hints} as introduction rules
 \lsti{dprolog} is not restricted to depth-first search, since
it follows the dictate of the given FPC; for example we could easily add
iterative-deepening search. Furthermore, the tactic can still use
\lsti{Hints} via Elpi's \lsti{Db} feature, and FPCs can provide a
\emph{trace} that may be more customizable than the one offered by 
\lsti{(e)auto}'s hard-wired \lsti{Debug} facility.

### 2021-03-04 papers/types2020/sec3.tex
3.3. talks abiyt conversion. More todos, a fix to fig 3, new file for dependent prolog tactic

\begin{figure}
\lstinputlisting[basicstyle=\ttfamily,language=lprolog,linerange={sigs-end}]{code/dep-kernel.sig}
\lstinputlisting[basicstyle=\ttfamily,language=lprolog,linerange={check-end}]{code/dep-kernel-v2.mod}
\caption{Implementation of a  dependent kernel for Coq terms.}
\label{fig:kernel}
\end{figure}

@@

The following
example shows how we can use the above tactic to do FPC-driven logic
programming in Coq and return a Coq proof-term:
\begin{lstlisting}
Inductive insert (x:nat) : list nat -> list nat -> Prop :=
| i_n : is_nat x -> insert x [] [x]
| i_s : forall y: nat, forall ys, x <= y -> insert x (y :: ys) (x :: y :: ys)
| i_c : forall y: nat, forall ys rs, y <= x -> insert x ys rs -> insert x (y :: ys) (x :: rs).
Lemma i1:  exists R, insert 2 [0;1] R.
eexists.
elpi dprolog height 10.
Qed.
Print i1.
ex_intro (fun R : list nat => insert 2 [0; 1] R) [0; 1; 2]
  (i_c 2 0 [1] [1; 2] (i_c 2 1 [] [2] (i_n 2)) : exists R : list nat, insert 2 [0; 1] R
\end{lstlisting}

The \lsti{dprolog} tactic can be seen as a programmable (and possibly
faster) version of Coq's \lsti{eauto}, although it currently lacks the
latter's capability of solving goals by
\lsti{reflexivity}; % as well as the modularity of using
% \lsti{Hints} as introduction rules
 \lsti{dprolog} is not restricted to depth-first search, since
it follows the dictate of the given FPC, for example we could easily add
iterative-deepening search. Furthermore, the tactic can still use
\lsti{Hints} via Elpi's \lsti{Db} feature, and FPCs can provide a
\emph{trace} that may be more customizable than the one offered by the
\lsti{(e)auto}'s hard-wired \lsti{Debug} facility.

>>>>

\begin{figure}
\lstinputlisting[basicstyle=\ttfamily,language=lprolog,linerange={sigs-end}]{../code/dep-kernel.sig}
\lstinputlisting[basicstyle=\ttfamily,language=lprolog,linerange={check-end}]{../code/dep-kernel-v2.mod}
\caption{Implementation of a  dependent kernel for Coq terms.}
\label{fig:kernel}
\end{figure}

@@

The following
example shows how we can use the above tactic to do FPC-driven logic
programming in Coq and return a Coq proof-term:
\begin{lstlisting}
Inductive insert (x:nat) : list nat -> list nat -> Prop :=
| i_n : insert x [] [x]
| i_s : forall y: nat, forall ys, x <= y -> insert x (y :: ys) (x :: y :: ys)
| i_c : forall y: nat, forall ys rs, y <= x -> insert x ys rs -> insert x (y :: ys) (x :: rs).
Lemma i1:  exists R, insert 2 [0;1] R.
eexists.
elpi dprolog height 10.
Qed.
Print i1.
ex_intro (fun R : list nat => insert 2 [0; 1] R) [0; 1; 2]
  (i_c 2 0 [1] [1; 2] (i_c 2 1 [] [2] (i_n 2)) : exists R : list nat, insert 2 [0; 1] R
\end{lstlisting}

The \lsti{dprolog} tactic can be seen as a programmable (and possibly
faster) version of Coq's \lsti{eauto}; % , although it currently lacks the
% latter's capability of solving goals by
% \lsti{reflexivity}; % as well as the modularity of using
% \lsti{Hints} as introduction rules
 \lsti{dprolog} is not restricted to depth-first search, since
it follows the dictates of the given FPC --- for example we could easily add
iterative-deepening search. Furthermore, the tactic can still use
\lsti{Hints} via Elpi's \lsti{Db} feature, and FPCs can provide a
\emph{trace} that may be more customizable than the one offered by 
\lsti{(e)auto}'s hard-wired \lsti{Debug} facility.
